SAT Competition 2020\tnoteref{title}

### 2021-01-28 paper/main.tex
updated plots
resized my figures

\subsection{Similarity of Solvers}
To investigate how similar the solvers of the Main track are to each other, we
define a similarity metric based on the measured runtimes. For that we first
remove $84$ benchmarks that have not been solved by any solver. For the rest of
the benchmarks we use the PAR-2 score,
% based on the \num{5000}-second timeout
i.\,e., for each solver we set a score of \num{10000} for unsolved instances.
This leaves us with $316$ floats
$S_{1},\dots,S_{n}$ for each solver $S$. The \emph{similarity} of two solvers $S$
and $S'$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculated the \emph{similarity} between the $30$ solvers which achieved the
best PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}. Additionally the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Thus the two solvers are joined low in the
dendrogram. A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

One big cluster we can identify are the \solver{Maple}-decedents to the right,
all of which are modifications of the winners of the SAT Competition 2018 and
the SAT Race 2019.
The similarity within the cluster is high, with the exception of
\solver{Scavel} and \solver{exp\_V\_MLD\_CBT\_DL}.
They form a subcluster
with a lower similarity to the rest, but a high similarity within. In fact the highest measured similarity,
 % of $0.96$
besides the aforementioned \solver{CaDiCaL-alluip}, is observed between them.
The two solvers are both based on
\solver{MapleLCMDistChronoBT-dl-v2.2} but have different authors.
The high similarity suggest that the
changes they made either result in very similar behavior or do not have a
significant impact on the runtime performance.

The two configurations of the \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same code base as a lot of the solvers in the Maple-Cluster but their
overall performance is better and closer to the CMS-Cluster.

Close to them are the three \solver{CMS} versions. They differ in their implementation of stochastic local search
(SLS) and have similar performance.
The 2020 version of \solver{CaDiCaL}
performed worse than the modifications based on the 2019 version and
does not quite fit into any cluster.

The last cluster to the left of the heat map is comprised of the other solvers
originally written by \emph{Biere}. Interesting to note is that the
\solver{Kissat} configuration specialized for unsatisfiable instances joins the
others in the cluster high in the dendrogram. In fact \solver{Kissat-unsat} has
the lowest average similarity to all other solvers in the top 30. This hints at
its importance for the use in an optimal portfolio.
% Average similarity
% | Kissat-unsat         |  21.75012381941168 |
% | Kissat               | 22.591060319234913 |
% | CaDiCaL-alluip       |  22.93588211785061 |
% | CaDiCaL-alluip-trail |  22.93631201417939 |
% | Kissat-sat           |  22.99738644487821 |

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=\linewidth]{similarity/cross.pdf}
    \caption{Heat-map and dendrogram (top) based on the runtime similarity of the solvers participating in the main track.
      Darker regions mean that the solvers are more similar.
      A more precise relation between color and similarity-value together with a histogram of the values that appear is given at the bottom.
    }
    \label{fig-similarity-main}
\end{figure}

@@

We use random sampling to select a subset of benchmarks used in the Main track.
We start with $316$ benchmarks that have been solved by at least one solver in
the Main track and remove a number of benchmarks randomly.

The solvers are assigned a new rank in ascending order of their PAR-2 score on
each random sample. Note that we never encounter a tie. This ranking can be
seen as an estimate of the original ranking.
If even relatively small random samples result in a good estimate, we can draw a
positive conclusion about the robustness of the ranking.
% We draw a positive conclusion about the robustness of the original ranking, if
% even the resulting estimate from relatively small random samples
% are good approximations.
To determine how similar the estimate is to the original ranking, we calculate
the \emph{Spearman's rank correlation coefficient} of the two rankings. For each
possible subset size (1--316) we generate 50 random samples.
% and compute the correlation coefficient.

Spearman's rank correlation coefficient
for two rankings $r_{1}$ and $r_{2}$
is defined by the following equation:
% \begin{equation}
% \label{eq-spearman}
\[
    \rho = 1 - \frac{6\cdot\sum_{s}{(r_{1}(s) - r_{2}(s))^2}}{n(n^2-1)},
\]
% \end{equation}
where $n$ is the number of solvers in the Main track ($48$), and the rankings
$r_{\{1,2\}}(s)$ map a solver $s$ to its rank, \emph{i.e.,} $1$ for the best
performing solver, whereas $48$ is assigned to the solver
with the highest average PAR-2 score.

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient: The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch all pairs of adjacent solvers in
rank still gives a value of $\rho = 0.9974$. On the other hand, switching the
highest ranked solver (\solver{Kissat-sat}) with the lowest ranked solver,
results in a rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to
the bottom of the ranking and moving every other solver up one rank gives a
higher $\rho = 0.8776$. Doing the same to all three \solver{Kissat}
configurations (with ranks 1, 2 and 11, now 46, 47 and 48) results in
$\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.
\begin{figure}[h]
  \label{fig:sampleAll}
  \includegraphics[width=\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances, 50 samples for each size}
\end{figure}
The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, still results in an average rank correlation above $0.96$

This suggest that the influence of the random selection in Algorithm~\ref{algo:select} on the solver ranking is limited.
% \ref{sec:byob}
The collected data cannot show, whether all of the benchmarks originally
submitted by the solver authors, have a systematic bias. However since each
newly submitted benchmark family originates from a different domain and is often
the result of current research, we can assume that the submitted families
together are representative.

\begin{figure}[h]
  \label{fig:sampleFamily}
  \includegraphics[width=\textwidth]{stability/fam_leave_one_out_corr.png}
  \caption{Change of rank correlation under removal of individual benchmark families}
\end{figure}

Figure~\ref{fig:sampleFamily} shows the rank
correlation coefficient resulting from removing a complete benchmark family
instead of a random subset. As expected removing a non random subset can have a
higher impact on the ranking even if it is small. Removing the 13 benchmarks
from the \emph{hgen}-family results in a (still high) rank correlation of
$0.9895$. Additionally, the ranking of the top five solvers stays the same.
The individual removal of all other benchmarks families results in a rank correlation over $0.99$.

>>>>

\subsection{Similarity of Solvers}
To investigate the similarity of solvers from the Main track, we define a
similarity metric based on measured runtimes. We start by removing $84$
benchmarks that have not been solved by any solver. For the rest, a PAR-2 score
is assigned to each instance for every solver,
% based on the \num{5000}-second timeout
\emph{i.e.}, we set a score of \num{10000} for unsolved instances.
Each solver $S$ thus is associated with the PAR-2 scores $S_{1},\dots,S_{316}$.
% This leaves us with $316$ floats
% $S_{1},\dots,S_{316}$ for each solver $S$.
The \emph{similarity} of two solvers $S$
and $S'$, normalized to the interval $[0,1]$ is defined
as:
% by Equation~\ref{eq-similarity}.

% \begin{equation}
% \label{eq-similarity}
\[
	similarity(S, S') =\quad 1 - \frac{\sum{|S_{i} - S'_{i}|}}{316 \cdot \num{10000}}
\]
% \end{equation}
We calculated the \emph{similarity} of the $30$ solvers with the
best average PAR-2 score in the main track. The results are displayed in
Figure~\ref{fig-similarity-main} as a \emph{heat map}, similar to the visualization in \cite{Balyo:2015:SATRACE}. Additionally, the result
of hierarchically clustering the solvers based on their similarity is
illustrated as a \emph{dendrogram}. The height at which two solvers or clusters
are joined reflects how similar they are. For example, enabling the \emph{trail
  saving} in \solver{CaDiCaL-alluip} has no impact on the runtime, resulting in
a similarity above \num{0.999}. Therefore the two solvers are joined low in the
dendrogram.
% A similar visualization was used for the results of the SAT Race 2015 \cite{Balyo:2015:SATRace}.

One big cluster that we identify consists of the \solver{Maple}-decedents,
all of which are modifications of the winners in the SAT Competition 2018 and
the SAT Race 2019.
The similarity within the cluster is high, with the exception of
\solver{Scavel} and \solver{exp\_V\_MLD\_CBT\_DL}.
They form a subcluster
with a lower similarity to the rest, but a high similarity within. In fact, the highest measured similarity,
 % of $0.96$
besides the aforementioned \solver{CaDiCaL-alluip}, is observed between them.
The two solvers are both based on
\solver{MapleLCMDistChronoBT-dl-v2.2} but have different authors.
The high similarity suggest that the
changes they made either result in a very similar behavior or do not have a
significant impact on the runtime performance.

The two configurations of \solver{Relaxed} by \emph{Zhang} and \emph{Cai}
use the same codebase as a lot of solvers in the Maple-Cluster but their
overall performance is better and closer to the \solver{CMS}-Cluster.
The three \solver{CMS} configurations
differ in their implementation of stochastic local search
(SLS) and have similar performance.
The 2020 version of \solver{CaDiCaL}
performed worse than the modifications based on the 2019 version and
does not quite fit into any cluster.

The last cluster to the left of the heat map is comprised of the other solvers
originally written by \emph{Biere}. Interesting to note is that the
\solver{Kissat} configuration specialized for unsatisfiable instances joins the
others in the cluster high in the dendrogram. In fact \solver{Kissat-unsat} has
the lowest average similarity to all other solvers in the top 30.
This suggests its importance for an optimal
portfolio.
% Average similarity
% | Kissat-unsat         |  21.75012381941168 |
% | Kissat               | 22.591060319234913 |
% | CaDiCaL-alluip       |  22.93588211785061 |
% | CaDiCaL-alluip-trail |  22.93631201417939 |
% | Kissat-sat           |  22.99738644487821 |

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{similarity/cross.pdf}
    \caption{Heat-map and dendrogram (top) based on the runtime similarity of the solvers participating in the main track.
      Darker regions mean that the solvers are more similar.
      A more precise relation between color and similarity-value together with a histogram of the values that appear is given at the bottom.
    }
    \label{fig-similarity-main}
\end{figure}

@@

We first use random sampling to select subsets of benchmarks used in the Main track.
We start with $316$ benchmarks that have been solved by at least one solver in
the Main track and remove a number of benchmarks randomly. For each
possible subset size (1--316) we generate 50 random samples.

The solvers are assigned a new rank in ascending order of their PAR-2 score on
each random sample. Note that we never encounter a tie. This ranking can be
seen as an estimate of the original ranking.
If even relatively small random samples result in a good estimate, we draw a
positive conclusion about the robustness of the ranking.
% We draw a positive conclusion about the robustness of the original ranking, if
% even the resulting estimate from relatively small random samples
% are good approximations.
To determine how similar an estimate is to the original ranking, we calculate
the \emph{Spearman's rank correlation coefficient} of the two rankings.
% and compute the correlation coefficient.

Spearman's rank correlation coefficient
for two rankings $r_{1}$ and $r_{2}$
is defined by the following equation:
% \begin{equation}
% \label{eq-spearman}
\[
    \rho = 1 - \frac{6\cdot\sum_{S}{(r_{1}(S) - r_{2}(S))^2}}{n(n^2-1)},
\]
% \end{equation}
where $n$ is the number of solvers in the Main track ($48$), and the rankings
$r_{\{1,2\}}(S)$ map a solver $S$ to its rank, \emph{i.e.,} $1$ for the best
performing solver, whereas $48$ is assigned to the solver
with the highest average PAR-2 score.

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient. The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch all pairs of adjacent solvers in
rank still gives a value of $\rho = 0.9974$. On the other hand, switching the
highest ranked solver (\solver{Kissat-sat}) with the lowest,
results in a rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to
the bottom of the ranking while moving every other solver up one rank gives a
higher $\rho = 0.8776$. Doing the same to all three \solver{Kissat}
configurations (with ranks 1, 2 and 11, now 46, 47 and 48) results in
$\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.
\begin{figure}[h]
  \label{fig:sampleAll}
  \centering
  \includegraphics[width=0.8\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances over 50 samples for each size}
\end{figure}
The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, still results in an average rank correlation above $0.96$.

This suggests that the impact of the random selection in Algorithm~\ref{algo:select} on the solver ranking is limited.
% \ref{sec:byob}
The collected data cannot show whether all of the benchmarks originally
submitted by the solver authors have a systematic bias. However, since each
newly submitted benchmark family originates from a different domain and is often
the result of current research, we can assume that the submitted families
together are representative.

\begin{figure}[h]
  \label{fig:sampleFamily}
  \centering
  \includegraphics[width=0.9\textwidth]{stability/fam_leave_one_out_corr.png}
  \caption{Change of rank correlation under removal of individual benchmark families}
\end{figure}

Figure~\ref{fig:sampleFamily} shows the rank
correlation coefficient resulting from removing a complete benchmark family
instead of a random subset. As expected, removing a non random subset can have a
higher impact on the ranking even if it is small. Removing the $13$ benchmarks
from the \emph{hgen}-family results in a (still high) rank correlation of
$0.9895$. Additionally, the ranking of the top five solvers stays the same.
The individual removal of all other benchmarks families results in a rank correlation above $0.99$.
